---
title: "Using lakinr"
author: "Lakin Robinson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r set_global_options, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Package Overview

`lakinr` package reads and plots Fatality Analysis Reporting System (FARS) data 
from the [National Highway Traffic Safety 
Administration](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars).

[FARS](https://en.wikipedia.org/wiki/Fatality_Analysis_Reporting_System) is a
nationwide census, providing the American public yearly data, regarding fatal
injuries suffered in motor vehicle traffic crashes.


## Load Packages

```{r}
library(lakinr)
library(maps)
```


## Included data

The following data files are included within the package:

```{r}
system.file('extdata', 'accident_2013.csv.bz2', package = 'lakinr')
system.file('extdata', 'accident_2014.csv.bz2', package = 'lakinr')
system.file('extdata', 'accident_2015.csv.bz2', package = 'lakinr')
```
These files are used later to demonstrate the use of the package's functions.

## Package Functions

### Function `make_filename`

This function returns file name using the relevant year. An example is shown below:

```{r}
year <- 2013
make_filename(year)
```

### Function `fars_read`

This function accesses `readr::read_csv` in otder to pars a FARS data file or, in the case of a nonexistent file, return an error. An example of each case is shown below:

```{R}
#file exists
data<-fars_read('accident_2013.csv.bz2')

#file does not exist
data2<-fars_read('accident_3000.csv.bz2')
```

### Function `fars_read_years`

This is a function used by `fars_summarize_years`. The purpose is to parse files that correspond with given year(s). An example is shown below:

```{r}
years<-2013:2015
fars_read_years(years)
```


### Function `fars_summarize_years`

This function constructs a summary of yearly accidents data, by month. It uses the functionality of fars_read_years to do so. An example is shown below:

```{r}
years<-2013:2015
fars_summarize_years(years)
```

### Function `fars_map_state`

The function plots the location of individual accideants for a given `year` on a map of a given state (`state_num`). If the `state_num` is invalid, an error is returned. If there are no accidents for the given state/year combination, a message is returned. Two examples are shown below:

```{r}
#State input is valid
state_num <- 49 # Utah
year <- 2014
fars_map_state(state_num, year)

#State input is invalid
state_num<-350
year<-2015
fars_map_state(state_num, year)
```


# Links:
* [Package lakinr GitHub]()
* [Package lakinr Travis CI]()
